### 2022-02-04 outputs/paper/roughpaper.Rmd
some legwork for paper

---
title: "Week 3 Tutorial"
author: "Jack McKay"
date: "1/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages("opendatatoronto")
library(opendatatoronto)
library(tidyverse)
library(tidyr)
library(knitr)
library(janitor)
library(lubridate)

citation("opendatatoronto")
```

```{r}
lct <- Sys.getlocale("LC_TIME"); Sys.setlocale("LC_TIME", "C")
Sys.setlocale("LC_TIME", lct)
```

```{r}
## Acquire Dataset ##
populations <-
  list_package_resources("ac77f532-f18b-427c-905c-4ae87ce69c93") %>% 
  filter(name == "toronto-shelter-system-flow")  %>%  
  get_resource()
```

```{r} 
populations$`date(mmm-yy)` <- Map(paste, populations$`date(mmm-yy)`, "-01", sep="")
populations$`date(mmm-yy)` <- strptime(populations$`date(mmm-yy)`, "%b-%y-%d")
populations$`date(mmm-yy)` <- as.Date(populations$`date(mmm-yy)`, format = "%Y-%m-%d")
names(populations)[2] <- "Date"
```

```{r}
## Clean Dataset ##
homeless_type <-
  populations %>% 
  select(population_group, Date, ageunder16, `age16-24`, `age25-44`, `age45-64`, age65over)

homeless_type <- 
  homeless_type %>% 
    tidyr::gather("Age Group", "Population", 3:7)

write_csv(
  x = homeless_type, 
  file = "homeless_types.csv"
  )
```

```{r} 
read_csv("cleaned_populations.csv")
populations_clean$`date(mmm-yy)` <- Map(paste, populations_clean$`date(mmm-yy)`, "-01", sep="")
populations_clean$`date(mmm-yy)` <- strptime(populations_clean$`date(mmm-yy)`, "%b-%y-%d")
populations_clean$`date(mmm-yy)` <- as.Date(populations_clean$`date(mmm-yy)`, format = "%Y-%m-%d")
names(populations_clean)[1] <- "Date"
```

```{r}
#### Explore ####

# This graph shows the breakdown of homeless populations, grouped by age, based on population group, on January 1st, 2020

homeless_pop <- 
  homeless_type %>% 
  filter(Date == as.Date("2021-01-01")) %>% 
  filter(population_group != "All Population")

homeless_pop %>% 
  ggplot( aes(x = `Age Group`, y = Population, fill = population_group, label = Population)) +
  geom_col(stat="identity") +
  geom_text(stat = "identity", size = 2, position = position_stack(vjust = 0.5), check_overlap = TRUE)
```

Here we can see that by far the most populous age group is 25-44, accounting for 40.42% of the total homeless population.

Additionally, we note that an especially high proportion of homeless people younger than 16 years of age are refugees; 35.37%. In comparison, of the 7210 homeless people 16 or older, only 776 are refugees (10.76%).

Conversely, Indigenous people comprise a remarkably low proportion of the under 16 age group, accounting for less than 1%, compared to 12.22% of the total population older than 16.

```{r}
## Explore ##

# This graph shows how homeless populations, grouped by age, fluctuate throughout the year

total_homeless <-
  homeless_type %>% 
  filter(population_group == 'All Population')

total_homeless %>% 
  ggplot(aes(x = Date, y = Population, color = `Age Group`, group = 2)) +
  scale_x_date(date_labels = "%Y-%m") + 
  geom_point() +
  geom_line(aes(group = `Age Group`))
```

Here we see that the homeless populations aged 45 and older have risen over the last 2 years, whereas the population younger than 45 has declined. Similarly, aside from the 65+ age group, we notice a substantial decrease in shelter flow across all populations between January 2020 and July 2020. This decrease is by far the most dramatic in the 25-44 age group. This could be because that is the age group that is most likely to be age to survive outside of shelters